Texture mapping 3D planar models of indoor environments with noisy camera poses

### 2014-01-04 spie2014/paperv2.tex
organize stuff

\documentclass[]{spie}  %>>> use for US letter paper
%%\documentclass[a4paper]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{float}
\title{Texture mapping 3D planar models of indoor environments with noisy camera poses}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}
  Automated 3D modeling of building interiors is used in applications
  such as virtual reality and environment mapping. Texturing these
  models allows for photo-realistic visualizations of the data
  collected by such modeling systems. We perform these tasks using a
  backpack-mounted data acquisition system. Camera poses obtained by
  this system and used for texturing often suffer from inaccuracies
  however, resulting in visible discontinuities when successive images
  are projected onto a surface for texturing. Existing methods to
  stitch images together are often computationally expensive and work
  independently of pose estimates and geometry information. By
  intelligently selecting images for texturing, we develop an
  efficient method to adjust camera poses in 2D, using linear features
  to match images to geometry, and point features to match images
  together. This is followed by two different methods to composite
  images together, based on the geometry of surfaces being
  textured. The effectiveness of our method is demonstrated on a
  number of different indoor environments.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction} % \label{} allows reference to this section
Three-dimensional modeling of indoor environments has a variety of
applications in fields such as architecture, conservation, and
simulation. Applying accurate textures to these models is important to
add context to digital visualizations, and allows for image-based
analysis of environments. In this paper, we perform data acquisition
with a backpack-mounted system carried by an ambulatory human
\cite{liu2010indoor}. This backpack system contains 2 fisheye cameras
as well as laser scanners and inertial measurement units,
used to solve the simultaneous localization and mapping (SLAM)
problem, which produces backpack (and therefore camera) poses over
time as well as a point cloud representing the surrounding environment
\cite{chen2010indoor, liu2010indoor, kua2012loopclosure}. SHOW PICTURE OF BACKPACK. For
environment geometry, we work with models obtained by fitting
low-resolution planar surfaces to these point clouds
\cite{sanchez2012point}.

While a human-carried system provides advantages over more common
wheeled systems in terms of agility and portability, it often leads to
relatively high localization error even after applying sophisticated
localization techniques \cite{liu2010indoor}. Furthermore, our model
geometry usually contains some error as well, and is often tuned to
ignore low-resolution features, such as small walls, furniture, and
other minor non-planar features commonly found in indoor
environments. As a result, methods for texture mapping that rely on
accurate camera poses and projection surfaces generally produce poor
results, while more sophisticated methods to accurately align images
in 3D lead to high runtime on our typically very large datasets. In
contrast, we present a method, which given a high number of images,
searches for a subset of them that can be successfully aligned with 2D
adjustments alone.

The remainder of the paper is organized as follows. Section
\ref{sec:simpleTextureMapping} explains how images are projected onto
our geometry and explains a simple approach towards selecting images
for texturing. Section \ref{sec:existingApproaches} covers existing
approaches to image stitching, and their performance on our
datasets. Section \ref{sec:2dAlignment} contains our approach towards
efficient 2D image alignment, followed by Section
\ref{sec:imageCompositing}, which describes our two methods of
compositing images. Section \ref{sec:resultsAndConclusions} contains
results and conclusions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The geometry of the texture mapping process for a planar surface is
shown in Figure \ref{fig:projection}.  We are provided with a set of
$M$ images to texture a target plane, where each image has a camera
matrix $P_i$ for $i=1..M$, which translates a 3D point in the world
coordinate system to a 2D point or pixel in image $i$'s coordinates. A
camera matrix $P_i$ is composed of the camera's intrinsic parameters,
containing focal length and image center, as well as extrinsic
parameters which specify the rotation and translation of the camera's
position in 3D world coordinates at the time that image $i$ is
taken. These extrinsic parameters are determined by the backpack
hardware and localization algorithms \cite{chen2010indoor,
  liu2010indoor, kua2012loopclosure} and are substantially noisy.

Because our backpack system takes photos at a rate of 5 Hz, hundreds
of images are available for texturing each surface in our model. Our
goal in designing a texture mapping process is to decide which of
these images should be used, and where their contents should map onto
the final texture, in order to minimize any visual discontinuities or
seams that would suggest that the plane's texture is not composed of a
single continuous image.

\begin{figure}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.8in]{Projection.pdf}
    \caption{Surfaces to be textured are specified in 3D space by
      corners $C_i$. Images are related to each surface through the
      camera matrices $P_{1..m}$. }
    \label{fig:projection}
  \end{minipage}
  \hspace{0.5cm}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.8in]{scoringFunction.jpg}
    \caption{We minimize camera angle $\alpha$ and distance $d$ by
      maximizing the scoring function $\frac{1}{d} (-1 \cdot \vec{c})
      \cdot \vec{n}$}
    \label{fig:scoringFunction}
  \end{minipage}
\end{figure}

\subsection{Tile-based Texture Mapping}
\label{sec:tileBasedMapping}
Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
a simple texturing approach can be performed by discretizing the
target plane into small square tiles, and choosing an image to texture
each tile directly.

We choose to work with rectangular units to ensure that borders
between any two distinct images in the final texture are either
horizontal or vertical. Since most strong environmental features
inside buildings are horizontal or vertical, any visible seams in our
texture will intersect them minimally and be less noticeable.

In order to select an image for texturing a tile $t$, we first gather
a list of candidate images that contain all four of its corners, which
we can quickly check by projecting $t$ into each image using the $P_i$
camera matrices. Furthermore, each candidate image must have been
taken at a time when its camera had a clear line-of-sight to $t$,
which can be calculated using standard ray-polygon intersection tests
between the camera location, $t$, and every other plane
\cite{rayintersection}.

Once we have a list of candidate images for $t$, we define a scoring
function in order to objectively select the best image for texturing
$t$. Since resolution decreases and camera pose errors become more
extreme over distance, we wish to minimize the distance between
cameras and the surfaces they texture. Additionally, we desire images
that are projected perpendicularly onto the plane, maximizing the
resolution and amount of useful texture available in their
projections, as well as minimizing any parallax effects due to
real-world geometry not accurately represented by our digital
model. In other words, we wish to minimize the angle between the
tile's normal vector and the camera axis for images selected for
texturing that tile. These two criteria can be met by maximizing the
function $\frac{1}{d} (-1 \cdot \vec{c}) \cdot \vec{n}$ as shown in
Figure \ref{fig:scoringFunction}. Specifically, $d$ is the distance
between the centers of a camera and a tile, and $\vec{n}$ and
$\vec{c}$ are the directions of the plane's normal and the camera axis
respectively.

As Figure \ref{fig:compareAll}(a) demonstrates, there are many image
boundaries with abrupt discontinuities between tiles, due to
significant misalignment between images, though most of it appears to
be reconcilable using 2D transforms. Given our high number of input
images, a better image selection procedure will further improve
results, as described in Section \ref{sec:imageCompositing}, but a
more significant and reliable improvement can be found through first
improving the alignment of our images.

\begin{figure}[H]
  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_naive.jpg}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_naive_shift.jpg}}

@@

@@

  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=3.4in]{wall2_cache_comparison.png}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=3.4in]{wall2_shortest_comparison.png}}

  \caption{(a): Tile-based texturing. (b): Tile-based texturing after
    image alignment. (c): Tile-based texturing after image alignment
    with caching. (d): Shortest path texturing after image
    alignment). (e,f): Blending applied to (c) and (d). (g,h): Zoomed
    in views of discontinuities in (e) vs. in (f).}
  \label{fig:compareAll}
\end{figure}

\section{Existing Approaches to Image Alignment}
\label{sec:existingApproaches}
Stitching together multiple images to produce a larger, seamless image
is a commonly performed task, with many successful approaches over the
past years. Generally, parts of images are matched to each other,
usually through direct comparisons or feature detection and
matching. Images are then transformed to maximize matches, often by
calculating homographies between pairs of images, or by iteratively
adjusting camera poses in 1 to 6 degrees of freedom.

Feature matching has a number of advantages over direct matching that
make it more suitable for our often non-planar data, and rotational
differences \cite{szeliski2006image}. Feature matching however, works
best when multiple unique visual references exist in the environment
that can be detected in multiple images. In contrast, our indoor
environments have a high prevalence of bare surfaces, as well as
repeating textures, such as similar windows, doors, and wall-mounted
decorations, that cause difficulty in disambiguating features. This
lack of reliable reference points often results in errors when
matching images together.

Additionally, our datasets often contain long chains of images, which
can lead to large overall error when images are matched together. For
example, a slight error in image correspondences in the middle of a
chain could lead to high global error at one or both ends of the
chain.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=6in]{graphApproach.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=6in]{autostitchResult.jpg}}

@@

  \caption{Texture alignment via (a) the graph-based localization
    refinement algorithm, (b) the AutoStitch software package, and (c)
    our method.}
  \label{fig:mosaic3D}
\end{figure}

Past work with our data acquisition backpack integrated image
stitching with an iterative localization algorithm, and performed
homography-based camera pose refinement \cite{liu2010indoor}. When run
on long chains of images, especially where features are sparse, it
produces distorted textures, as seen in Figure \ref{fig:mosaic3D}(a),
due to the problems described above. This approach is also not
closed-form, and its iterative camera adjustment process over our
large datasets led to prohibitively long computation time.

The AutoStitch software package performs homography-based alignment as
well, with additional provisions that attempt to reduce drift and
increase efficiency \cite{panorama2d, autostitch}. Though AutoStitch
also performs well in areas with dense features, it can not handle
areas without features, and has trouble aligning wall sections with
even short segments of bare texture. The example in Figure
\ref{fig:mosaic3D}(b) was generated after manual tuning, and areas
with fewer visual features or repeating texture patterns simply failed
outright with AutoStitch.

More importantly, neither approach

\section{2d Image Alignment}
\label{sec:2dAlignment}

In this section, we describe our method of efficient and robust image
alignment. Though state-of-the-art techniques towards image stitching
work in 3D, we elect to perform image alignments entirely in
2D. Besides the benefits of reduced complexity, we have found 2D
alignment to work well for two main reasons. First, because we have
such a large number of images to choose from, the compositing
approaches in Section \ref{sec:imageCompositing} are able to select a
subset of images that align well, given only 2D alignments. Second,
the nature of our input data is such that localization error chiefly
occurs in two dimensions, in the same plane as the surface being
projected onto. Our data acquisition system contains cameras facing to
the sides of the operator. The operator makes efforts to walk parallel
to walls, and the localization and model-generation algorithms that
provide our input conform the operator's path to be straight and
parallel to detected surfaces \cite{kua2012loopclosure,
  sanchez2012point}. Since the operator also walks in a manner to keep
the backpack balanced, errors in roll and yaw are minimal. Thus, our
highest errors stem from uncertainty in the operator's pitch,
equivalent to rotation around the camera axes, as well as location
along surfaces. These equate to 2D rotation and translation in a
surface's plane.

Our 2D alignment procedure consists of three parts. First, all images
are projected onto the surface and lines within these projections are
detected. These lines are then rotated and shifted to match
geometry-based lines comprising the surface's boundary and
intersection with other surfaces. Second, occlusion checks are
performed to remove invalid parts of each image for the target
surface. These two steps are image-independent and performed in
parallel. Third, we detect SIFT feature matches between pairs of
images and solve a weighted linear least squares problem in 2D to
maximize matches.

\subsection{Geometry-based Alignment}
\label{sec:geometryAlignment}
After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we detect line
segments in the image projections using Hough transforms, as shown in
Figure \ref{fig:geometryAlignment}(a). Experience and intuition show
that walls in indoor environments often contain linear features that
are either horizontal or vertical, corresponding to doors, windows,
posters, etc. Thus, when texturing walls, we rotate images in 2D such
that dominant lines are made to be horizontal or vertical. This can be
further generalized by instead orienting lines to be parallel or
perpendicular with a surface's boundaries, which makes it applicable
to ceilings and floors, where inner walls and interior features tend
to be parallel or perpendicular to outer walls.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[height=3in]{detectedLines.png}}
  \hspace{0.4 cm} \subfloat[][]{\includegraphics[height=3in,
    width=4in]{bothunalignedaligned.png}}
  \caption{(a) Lines detected in image projections via Hough
    transform.  (b) In this example, we project two images at once for
    the sake of demonstration. Red line segments represent where
    surfaces intersect, and green line segments are line segments
    detected in the image. Above are the original projections using
    input camera poses, and below are the projections after rotation
    and translation for alignment.}
  \label{fig:geometryAlignment}
\end{figure}

Furthermore, at this point in time, image occlusions have not been
accounted for. As a result, some image projections contain texture
that should project to an adjacent surface, sometimes with a visible
linear boundary where the two surfaces meet. If this linear boundary
can be detected by Hough transform, the image is rotated and shifted
in 2D such that the visual boundary between two surfaces in an image
projection matches the physical boundary in our digital model. An
example of such an adjustment is in Figure
\ref{fig:geometryAlignment}(b) and \ref{fig:geometryAlignment}(c).

To perform these alignments, we use a RANSAC \cite{fischler1981random}
framework to determine an optimal rotation angle between lines in our
image and lines in our geometry. RANSAC allows us to ignore erroneous
lines in our image, such as the slanted lines corresponding to the
ladder in Figure \ref{fig:geometryAlignment}(a). Once this rotation is
discovered and applied, we gather all pairs of image line segments and
geometry line segments with less than $0.1^\circ$ difference in
orientation and less than 250 mm distance at their furthest points. If
2 or more pairs exist, we select the 2 that contain the longest
noncollinear image-based line segments, and perform the corresponding
fixed translation to match them. If we only have 1 pair, we perform
the minimal shift to match the lines of that pair, and make note of
their orientation, for usage in Section
\ref{sec:robustSIFTFeatureMatching}.

\subsection{Image Occlusion}
\label{sec:imageOcclusion}
Now that images have been aligned to geometry where possible, we
perform a simple recursive occlusion procedure on each image. This is
done by performing the intersection tests in Section
\ref{sec:simpleTextureMapping} in a regularly spaced grid. Where four
corners of a rectangular region are occluded, texture is
removed. Where no corners are occluded, the recursion stops. Where
there is a mixture of both, the rectangular region is subdivided into
four, and the same process is performed on each. By performing Section
\ref{sec:geometryAlignment}'s alignment procedure before occlusion,
texture belonging to other surfaces is accurately removed, which is
necessary for the next section.

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align images by searching for corresponding points
between all pairs of overlapping images. We use feature alignment
rather than pixel or intensity-based alignment due to the differences
in lighting as well as possible occlusion among our images, both of
which feature alignment is less sensitive to \cite{lowe1999object,
  mikolajczyk2005performance, szeliski2006image}.  We use SiftGPU
\cite{siftgpu} for its high performance on both feature detection as
well as pairwise matching. These matches determine $dx$ and $dy$
distances between each pair of features for two image projections,
though these distances may not always be the same for different
features. Since indoor environments often contain repetitive features
such as floor tiles or doors, we need to ensure that SIFT-based
distances are reliable. First, we only perform alignment on parts of
images that overlap given the original noisy poses. Second, we discard
feature matches that correspond to an image distance greater than 200
mm from what the noisy poses estimate. In order to utilize the
remaining feature matches robustly, RANSAC \cite{fischler1981random}
is again used to estimate the optimal $dx_{i,j}$ and $dy_{i,j}$
distances between two images $i$ and $j$. We use a 5 mm threshold for
RANSAC, so that SIFT matches are labeled as outliers if their distance
is not within 5 mm of the sampled average distance.

We now use the $dx_{i,j}$ and $dy_{i,j}$ distances between each pair
of images to refine their positions using weighted linear least
squares. The variables we wish to solve for are the $_ix$ and $y_i$
positions of our images, while our equations are the SIFT-based
distances between pairs of images, images fixed to geometry with 0 or
1 degrees of freedom, and the original camera poses. Ignoring
weighting and the original camera poses for now, an example setup for
solving $\textrm{min}_{\vec{\beta}} ||W^\frac{1}{2}(A \vec{\beta} -
\vec{\gamma})||_2^2 $ with 3 images is shown below.

>>>>

\documentclass[]{spie}  %>>> use for US letter paper
%%\documentclass[a4paper]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}

\title{Texture mapping 3D planar models of indoor environments with noisy camera poses}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}

  Automated 3D modeling of building interiors is used in applications
  such as virtual reality and environment mapping. Texturing these
  models allows for photorealistic visualizations of the data
  collected by such modeling systems. Camera poses obtained by these
  systems and used for texturing often suffer from inaccuracies
  however, resulting in visible discontinuities when successive images
  are projected onto a surface for texturing. Existing methods to
  stitch images together are often computationally expensive and work
  independently of pose estimates and geometry information. We propose
  an efficient method to refine camera poses in 2D using both existing
  estimates and geometry information, followed by two different
  methods to composite images together, based on the uniformity of
  available images. The effectiveness of our methods is demonstrated
  on a number of different indoor environments.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction} % \label{} allows reference to this section

Three-dimensional modeling of indoor environments has a variety of
applications such as training and simulation for disaster management,
virtual heritage conservation, and mapping of hazardous sites. Manual
construction of these digital models can be time-consuming, and as a
result, automated 3D site modeling has garnered much interest in
recent years.

The first step in automated 3D modeling is the physical scanning of an
environment's geometry. An indoor modeling system must recover its own
poses within an environment while simultaneously reconstructing the 3D
structure of the environment itself \cite{chen2010indoor,
  liu2010indoor, kua2012loopclosure}. This is known as the
simultaneous localization and mapping (SLAM) problem, and is generally
solved by taking readings from laser range scanners, cameras, and
inertial measurement units (IMUs) at multiple locations within the
environment.

In this paper, we work with data obtained from a backpack-mounted
system, carried by an ambulatory human. Such a system provides
advantages over more common wheel-mounted systems in terms of agility
and portability, but results in higher difficulty and error when
performing localization \cite{liu2010indoor}. As a result, common
approaches to texture mapping that rely on accurate camera poses
produce poor results, leading to the development of the approaches
contained in this paper. Besides localization information and images,
environment geometry is required as well for texture mapping. In this
paper we work with models obtained by fitting planar surfaces to point
clouds generated by the backpack system \cite{sanchez2012point}.

Though our texture mapping procedure was designed with the
aforementioned system in mind, it is applicable to other systems,
provided the required inputs of planar geometry, images with rough
extrinsic pose estimates, and either known or estimated intrinsic
camera calibration matrices. An important benefit of our method is its
relatively low complexity and runtime, and independence of many steps,
allowing for efficient tuning of parameters and easy accomodation of
new input data given updates in localization or model-generation
algorithms.

The remainder of the paper is organized as follows. Section
\ref{sec:simpleTextureMapping} explains how input images are projected
onto our geometry and a simple approach towards texturing. Section
\ref{sec:existingApproaches} covers more sophisticated existing
approaches, using image matching and stitching, and their performance
on our datasets. Section \ref{sec:2dAlignment} contains our approach
towards image alignment, followed by Section
\ref{sec:imageCompositing}, which describes our two methods of
compositing images. Section \ref{sec:resultsAndConclusions} contains
results and conclusions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The geometry of the texture mapping process for a planar surface is
shown in Figure \ref{fig:projection}.  As described earlier, we are
provided with a set of $M$ images to texture a target plane. Each
image has a camera matrix $P_i$ for $i=1..M$, which translates a 3D
point in the world coordinate system to a 2D point or pixel in image
$i$'s coordinates. A camera matrix $P_i$ is composed of the camera's
intrinsic parameters, containing focal length and image center, as
well as extrinsic parameters which specify the rotation and
translation of the camera's position in 3D world coordinates at the
time that image $i$ is taken. These extrinsic parameters are
determined by the backpack hardware and localization algorithms
\cite{chen2010indoor, liu2010indoor, kua2012loopclosure} and are quite
noisy.

Because our backpack system takes photos at a rate of 5 Hz, thousands
of images are available for texturing each surface in our model. Our
goal in designing a texture mapping process is to decide which of
these images should be used, and where their contents should map onto
the final texture, in order to minimize any visual discontinuities or
seams that would suggest that the plane's texture is not composed of a
single continuous image.

\begin{figure}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.5in]{Projection.pdf}
    \caption{Planes are specified in 3D space by four corners $C_1$ to
      $C_4$. Images are related to each plane through the camera
      matrics $P_{1..m}$. }
    \label{fig:projection}
  \end{minipage}
  \hspace{0.5cm}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.125in]{scoringFunction.jpg}
    \caption{We minimize camera angle $\alpha$ and distance $d$ by
      maximizing the scoring function $\frac{1}{d} (-1 \cdot \vec{c})
      \cdot \vec{n}$}
    \label{fig:scoringFunction}
  \end{minipage}
\end{figure}

\subsection{Tile-based Texture Mapping}
\label{sec:tileBasedMapping}
Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
a simple texturing approach can be performed by discretizing the
target plane into small square tiles, in our case 5 pixels across, and
choosing an image to texture each tile.

We choose to work with rectangular units to ensure that borders
between any two distinct images in the final texture are either
horizontal or vertical. Since most strong environmental features
inside buildings are horizontal or vertical, any visible seams in our
texture thus intersect them minimally and are less noticeable.

In order to select an image for texturing a tile $t$, we must first
gather a list of candidate images that contain all four of its
corners, which we can quickly check by projecting $t$ into each image
using the $P_i$ camera matrices. Furthermore, each candidate image
must have been taken at a time when its camera had a clear
line-of-sight to $t$, which can be calculated using standard
ray-polygon intersection tests between the camera location, $t$, and
every other plane \cite{rayintersection}.

Once we have a list of candidate images for $t$, we define a scoring
function in order to objectively select the best image for texturing
$t$. Since resolution decreases and camera pose errors compound over
distance, we wish to minimize the distance between cameras and the
surfaces they texture. Additionally, we desire images that are
projected perpendicularly onto the plane, maximizing the resolution
and amount of useful texture available in their projections, as well
as minimizing any parallax effects due to real-world geometry not
accurately represented by our digital model. In other words, we wish
to minimize the angle between the tile's normal vector and the camera
axis for images selected for texture mapping. These two criteria can
be met by maximizing the function $\frac{1}{d} (-1 \cdot \vec{c})
\cdot \vec{n}$ as shown in Figure
\ref{fig:scoringFunction}. Specifically, $d$ is the distance between
the centers of a camera and a tile, and $\vec{n}$ and $\vec{c}$ are
the directions of the plane's normal and the camera axis respectively.

\begin{figure}
  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_naive.jpg}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_naive_shift.jpg}}

@@

  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_cache_shift_details.png}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_shortest_shift_details.png}}

@@

  \caption{(a): Tile-based texturing. (b): Tile-based texturing after
    image alignment. (c): Tile-based texturing after image alignment
    with caching. (d): Shortest path texturing after image
    alignment). (e,f): Comparison of image artifacts in (c)
    vs. (d). (g,h): Blending applied to (c) and (d).}
  \label{fig:compareAll}
\end{figure}

As Figure \ref{fig:compareAll}(a) demonstrates, this approach leads to
the best texture for each tile independently, but results in many
image boundaries with abrupt discontinuities between tiles, due to
significant misalignment between images. Given our high number of
input images, a better image selection procedure could be used, as
described in Section \ref{sec:imageCompositing}, but a more
significant and reliable improvement can be found through the
alignment of our images.

\section{Existing Approaches to Image Alignment}
\label{sec:existingApproaches}
Stitching together multiple images to produce a larger image is a
commonly performed task, with many successful approaches over the past
few decades. Parts of images are first matched to eachother, through
direct pixel comparisons, or more commonly through feature detection
and matching. Images are then adjusted to maximize matches, either by
calculating homographies between pairs of images, or by modifying
camera poses in 1 to 6 degrees of freedom.

Feature detection and matching works best when multiple unique visual
references exist in the environment and are present within multiple
images. In contrast, our indoor environments have a high prevalence of
bare surfaces, as well as repeating textures that cause difficulty in
disambiguating features. This lack of reliable reference points often
results in errors when matching images together.

Additionally, our datasets often contain long chains of images, which
leads to error accumulation when image correspondences are not
accurate. For example, when matching a long chain of images through
homographies, a pixel in the $nth$ image is translated into the first
image's coordinates by multiplying by the $3\times3$ matrices $H_1 H_2
H_3 ... H_n$. Errors in any of these homography matrices are
propagated to all further images, resulting in drift. Figure
\ref{fig:mosaic3D}(a) shows the output of the AutoStitch software
package, which performs homography-based image mosaicing
\cite{autostitch}. Though AutoStitch performs well in areas with dense
features, it has difficulty with even short segments of bare walls,
and produces errors that cause straight lines to appear
distorted. Many areas with fewer visual features or repeating texture
patterns failed outright using AutoStitch.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=6in]{autostitchResult.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=6in]{graphApproach.jpg}}

@@

  \caption{Texture alignment via (a) image mosaicing, (b) the
    graph-based localization refinement algorithm from
    \cite{chen2010indoor}, and (c) the method proposed in this paper.}
  \label{fig:mosaic3D}
\end{figure}

Image projections can also be aligned by iteratively adjusting camera
poses to maximize matches. This is generally done over 6 degrees of
freedom \cite{liu2010indoor}. The approach in [2] works well locally,
but over larger areas results again in error accumulation and drift,
as shown in Figure \ref{fig:mosaic3D}(b). Furthermore, it adjusts
image locations using a long iterative process, which results in high
runtime.

\section{2D Image Alignment with Geometry Information}
\label{sec:2dAlignment}
In this section, we describe our method of efficient and robust image
alignment. Our approach consists of three main parts. First, all
images are projected onto the surface and lines within these
projections are detected. These lines are then matched up with
geometry-based lines comprising the surface's boundary and
intersection with other surfaces. Second, occlusion checks are
performed to remove invalid parts of each image for the target
surface. These two steps are image-independent and performed in
parallel. Third, we detect SIFT feature matches between pairs of
images and solve a linear least squares problem in 2D to maximize
matches.

All of our following calculations and alignments are performed in 2D,
partly for efficiency, partly because we have a large number of images
to choose from, and partly because the nature of our input data is
such that localization error chiefly occurs in two dimensions, in the
same plane as the surface being projected onto. Recall that our data
acquisition system is backpack-mounted, with cameras facing to the
sides of the operator. The operator makes efforts to walk parallel to
walls, and the localization and model-generation algorithms that
provide our input conform the operator's path to be straight and
parallel to detected surfaces \cite{kua2012loopclosure,
  sanchez2012point}. Since the operator walks in as upright a position
as possible, errors in roll, yaw, and the operator's distance from
parallel surfaces are minimal. Thus, our highest errors stem from
uncertainty in the operator's pitch, equivalent to rotation around the
camera axes, as well as location along surfaces. These equate to 2D
rotation and translation in a surface's plane.

\subsection{Geometry-based Alignment}
\label{sec:geometryAlignment}
After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we detect line
segments in the image projections using Hough transforms. Experience
and intuition show that walls in indoor environments often contain
linear features that are either horizontal or vertical, corresponding
to doors, windows, posters, etc. Thus, when texturing walls, we rotate
images in 2D such that dominant lines are made to be horizontal or
vertical. This can be further extrapolated by orienting lines with a
wall's boundaries, for example in areas with a slanted roof.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[height=1.5in]{geometryAlign_planes.png}}
  \subfloat[][]{\includegraphics[height=1.5in]{geomAlign}}
  \caption{(a) When texturing the red surface, it makes sense to align
    images to surface boundaries and intersections with other
    surfaces. (b) The darker lines are geometry-based lines, while
    lighter lines are lines detected in the image via Hough
    transform. Above is the original projection using the input camera
    poses, and below is the projection after rotation and translation
    for alignment.}
  \label{fig:geometryAlignment}
\end{figure}

At this point in time, image occlusions have not been accounted
for. As a result, some image projections contain texture that should
project to an adjacent surface, generally with a linear boundary where
the two surfaces meet. If this linear boundary is detected by Hough
transform, the image is rotated and shifted in 2D such that the visual
boundary between two surfaces in an image projection matches the
physical boundary in our digital model. An example of such an
adjustment is in Figure \ref{fig:geometryAlignment}

To perform the rotation and alignment, we use the RANSAC
\cite{fischler1981random} framework. RANSAC allows us to determine an
optimal rotation and translation to match up two groups of line
segments, while accounting for outliers. First, a rotation angle is
computed by gathering all pairs of image line segments and geometry
line segments with less than 20 degrees between them. We then run the
RANSAC algorithm by sampling 4 of these angles at a time, and using
their average as the fitting function. Angles over 2 degrees from this
average are considered outliers. Once this 2d rotation angle is
determined, it is applied to the image, and projections are
re-calculated and image-based lines are re-detected. Now, all pairs of
image line segments and geometry line segments with less than 0.2
degree difference and less than 250 mm distance at their furthest
points are gathered. RANSAC is then similarly employed to calculate
optimal horizontal and vertical translations independently, by
sampling 4 pairs at a time and considering an outlier threshold of 10
mm. If a translation is applied, we indicate that the image was
aligned to geometry in one or both dimensions, which will reduce
further shifting in Section \ref{sec:robustSIFTFeatureMatching}.

\subsection{Image Occlusion}
\label{sec:imageOcclusion}
Now that images have been aligned to geometry where possible, we
perform a simple recursive occlusion procedure on each image. This is
done by performing the intersection tests in section
\ref{sec:simpleTextureMapping} in a regularly spaced grid. Where four
corners of a rectangular region are occluded, texture is
removed. Where no corners are occluded, nothing occurs. Where there is
a mixture of both, the rectangular region is subdivided into four, and
the same process is performed on each. By performing Section
\ref{sec:geometryAlignment}'s alignment procedure before occlusion,
texture belonging to other surfaces is accurately removed, which is
necessary for the next section.

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images. We use
SIFT features for their high detection rate, and choose to use feature
alignment rather than pixel or intensity-based alignment due to the
differences in lighting as well as possible occlusion among our
images, both of which feature alignment is less sensitive to
\cite{lowe1999object, mikolajczyk2005performance, szeliski2006image}.
We use SiftGPU \cite{siftgpu} for its high performance on both feature
detection as well as pairwise matching. These matches determine $d^x$
and $d^y$ distances between each pair of features for two image
projections, though these distances may not always be the same for
different features. Since indoor environments often contain repetitive
features such as floor tiles or doors, we need to ensure that
SIFT-based distances are reliable. First, we only perform alignment on
parts of images that overlap given the original noisy poses. Second,
we discard feature matches that correspond to an image distance
greater than 40 pixels from what the noisy poses estimate. In order to
utilize the remaining feature matches robustly, RANSAC
\cite{fischler1981random} is again used to estimate the optimal
$d^x_{i,j}$ and $d^y_{i,j}$ distances between two images $i$ and
$j$. For this application, the RANSAC fitting function finds the
average distance between selected matches in a pair of images, and the
distance function for a pair of points is chosen to be the difference
between those points' SIFT match distance and the average distance
computed by the fitting function. We use a 10 pixel threshold, so that
SIFT matches are labeled as outliers if their horizontal or vertical
distances are not within 10 pixels of the average distance computed by
the fitting function.

We now use the RANSAC-calculated $d^x_{i,j}$ and $d^y_{i,j}$ distances
between each pair of images to refine their positions using weighted
linear least squares. There are a total of $M^{2}$ possible pairs of
images, though we only generate distances between images with SIFT
matches. Given these distances and the original image location
estimates, we can solve a least squares problem
($\textrm{min}_{\vec{\beta}} ||A \vec{\beta} - \vec{\gamma}||_2^2 $)
to estimate the location of the images on the plane. The
$M$-dimensional vector $\vec{\beta}$ represents the unknown $x$
location of each image on the plane for $1 \dots M$. The optimal $x$
and $y$ locations are obtained in the same way, so we only consider
the $x$ locations here:

\[\vec{\beta} =
\begin{pmatrix}
  x_1, & x_2, & x_3, & \cdots & x_{M-1}, & x_M
\end{pmatrix}
\]
